\documentclass[12pt,twoside,a4paper]{report}

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage[normalem]{ulem}
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}

\usepackage{multirow}
\usepackage{ctable}
\usepackage{colortbl}
\usepackage{subfig}
\usepackage{dcolumn}
\usepackage{url}


\newcommand{\myspace}{5cm}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{url}

% ADDED
\usepackage{ctable}
\usepackage{subfig}
\usepackage{array}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage[shortcuts]{extdash}

% nested calls
\renewcommand*\Call[2]{\textproc{#1}(#2)}

% NEW COMMANDS
\newcommand{\otoprule}{\midrule[\heavyrulewidth]}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% Function names
\DeclareMathOperator{\Trust}{Trust}
\DeclareMathOperator{\Aggregate}{Aggregate}
\DeclareMathOperator{\Median}{Median}
\DeclareMathOperator{\mor}{mor}
\DeclareMathOperator{\MOR}{MOR}
\DeclareMathOperator{\Rank}{\textsc{Rank}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

% elements
\newcommand{\D}{\mathit{D}}
\newcommand{\PD}{\mathit{PD}}
\newcommand{\I}{\mathit{I}}
\newcommand{\PT}{\mathit{PT}}
\newcommand{\T}{\mathit{T}}

% algs
\floatname{algorithm}{Listing}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\pushcode}[1][1]{\hskip\dimexpr#1\algorithmicindent\relax} % for condition over two lines

\newcommand{\nodewatcher}{\textit{nodewatcher}}
\newcommand{\wlanslovenija}{\textit{wlan slovenija}}

\frenchspacing
\begin{document}%\maketitle
\pagestyle{empty}
\begin{center} {\Large{Authors' response to the review of the paper COMNET-D-15-148}}\\
    ``\emph{\nodewatcher{}: A Substrate for Growing Your own Community Network}''\\
    submitted to Computer Networks
\end{center}

\noindent We would like to thank the Editor and the reviewers for their comments. We have found them to be very constructive and they have helped us greatly to improve the manuscript. The following list represents the changes made.
\begin{itemize}
\item Added some more references to relevant papers as suggested by the reviewers.
\item Improved text that describes Table 1 so that the used acronyms are also used in the text.
\item Fixed terminology in Figure 6 so that it is more consistent with the rest of the text.
\item In Section 4.5, under \textit{single installation} approach to federation, highlighted that for really large networks certain optimizations would still be necessary.
\item Removed some less relevant details of the firmware build process in Section 3.3.2.
\end{itemize}

\noindent In addition to these major changes (which are in detail explained in the subsequent pages as responses to concrete comments from the reviewers), we have also made several minor changes (corrected spelling mistakes, shortened some paragraphs, ...) that we do not highlight in this revision letter. \\

\noindent Sincerely,\\
Jernej Kos, Mitar Milutinović and Luka Čehovin

\end{document}
